---
title: "EDS 222 Week 3 Tuesday (Lab)"
subtitle: "Simulating data following a linear regression model"
format: html
---

```{r}
library(tidyverse)
set.seed(222)
theme_set(theme_bw())
```

Simulate data following a linear regression model.

Example: the relationship between water temperature and the length of copepods.

1. Choose parameters
2. Simulate data
3. Fit the model
4. Intepret the model

## Choose parameters

Our model has three parameters: two betas and a sigma. Our simulation needs one more parameter, our sample size.

```{r}
beta_0 <- 3000 # intercept: length of a copepod (in um) when temp is 0 degC
beta_1 <- -55 # slope: change in length per degC
sigma <- 100 # standard deviation of the residuals, the "spread" of the points around the line
n <- 50 # sample size
```

## Simulate data

1. Simulate the _predictor_
2. Calculate the _expected_
3. Simulate the _response_
4. Assemble a dataframe

### Simulate the predictor

For temperature _uniformly_ distributed random temps between 2 and 14 degC

```{r}
temp_c <- runif(n, min = 2, max = 14)
```

### Calculate the expected

```{r}
expected_length_um <- beta_0 + beta_1 * temp_c
```

### Simulate the response
_
The response is _normally_ distributed around the line

```{r}
length_um <- rnorm(n, mean = expected_length_um, sd = sigma) 
# notice the mean is the _vector_ of expected values
```

### Assembling the data frame

```{r}
# create a tibble
copepods <- tibble(temp_c, length_um)

# visualize the data
ggplot(copepods, aes(temp_c, length_um)) +
    geom_point()
```

## Fit the model

Use the `lm()` to fit a model. Specify the relationship between response and predictor with a _formula_.

```{r}
# length_um is responsing to temp_c
copepod_lm <- lm(length_um ~ temp_c, data = copepods)
```

Done and done!

## Interpret the model

Inspect the coefficients using `summary()`

```{r}
summary(copepod_lm)
```

Use `sigma()` to get the model estimate of sigma

```{r}
sigma(copepod_lm)
```

## Recap

When simulating data from a linear regression model:

1. Choose our parameters (betas, sigma, n)
2. Simulate our data (predictor, expected, response)
3. Fit the model (lm())
4. Interpret the model (summary(), sigma())