---
title: "Lab 2: Investigating claims using inference"
subtitle: "Permutation and bootstrapping"
format: html
---

# Background:

See [EDS22 Lab 2](https://eds-222-stats-f26.github.io/course-materials/labs/lab2.html).

**Claim 1:** Lead is elevated in the drip line compared to the rest of the yard.
- Question Type: Yes/no?
- Statistical Experiment: Permutation
**Claim 2:** Neighborhoods north of the Exide plant were exposed to more lead than those south.
- Question Type: How much?
- Statistical Experiment: Bootstrapping

```{r}
library(tidyverse)
set.seed(222)
```

```{r}
# Load data from /courses/EDS222
soil <- read_csv("/courses/EDS222/data/exide-soil-cores.csv")
soil
```

# Claim 1: Lead is elevated in the drip line compared to the rest of the yard.

## Hypotheses:

**Null Hypothesis (H~0~):** There is no difference in lead (Pb) concentration in the drip line compared to the rest of the yard.

**Alternative Hypothesis (H~A~):** The lead (Pb) concentration in the drip line is greater than the rest of the yard.

## Calculate observed value of the statistic:

```{r}
# Group Pb concentrations by location_code (DL or YD), calculate grouped means
soil_loc_means <- soil |> 
    summarize(pb_ppm_mean = mean(pb_ppm),
                .by = location_code)

# Calculate diff_means (DL - YD)
diff_means = soil_loc_means$pb_ppm_mean[1] - soil_loc_means$pb_ppm_mean[2]
```

## Shuffle the predictor variable and recalculate the statistic

```{r}
# Shuffle predictor variable (location_code)
shuffle_soil <- soil |> 
    mutate(shuffle_location_code = sample(location_code)) |>
    summarize(pb_ppm_mean = mean(pb_ppm),
                .by = shuffle_location_code)

# Calculate diff_means (DL - YD)
shuffle_diff_means <- shuffle_soil$pb_ppm_mean[1] - shuffle_soil$pb_ppm_mean[2]

```

## Repeat Shuffling

```{r}
# Replicate above shuffle 10K times (1e4)
shuffled_means <- replicate(1e4,{
                    # Shuffle predictor variable (location_code)
                    shuffle_soil <- soil |> 
                        mutate(shuffle_location_code = sample(location_code)) |>
                        summarize(pb_ppm_mean = mean(pb_ppm),
                                    .by = shuffle_location_code)

                    # Calculate diff_means (DL - YD)
                    shuffle_soil$pb_ppm_mean[1] - shuffle_soil$pb_ppm_mean[2]
                    })

#shuffled_means
```

```{r}
# Plotting histogram (null distribution)
tibble(shuffled_means) |> 
    ggplot(aes(shuffled_means)) + 
    geom_histogram()

```

## Compare the observed to the null

```{r}


```

## Interpret your results