---
title: "Lab 2: Investigating claims using inference"
subtitle: "Permutation and bootstrapping"
format: html
---

# Background:

See [EDS22 Lab 2](https://eds-222-stats-f26.github.io/course-materials/labs/lab2.html).

**Claim 1:** Lead is elevated in the drip line compared to the rest of the yard.
- Question Type: Yes/no?
- Statistical Experiment: Permutation
**Claim 2:** Neighborhoods north of the Exide plant were exposed to more lead than those south.
- Question Type: How much?
- Statistical Experiment: Bootstrapping

```{r}
library(tidyverse)
set.seed(222)
```

```{r}
# Load data from /courses/EDS222
soil <- read_csv("/courses/EDS222/data/exide-soil-cores.csv")
soil
```

# Claim 1: Lead is elevated in the drip line compared to the rest of the yard.

## Hypotheses:

**Null Hypothesis (H~0~):** There is no difference in lead (Pb) concentration in the drip line compared to the rest of the yard.

**Alternative Hypothesis (H~A~):** The lead (Pb) concentration in the drip line is greater than the rest of the yard.

## Calculate observed value of the statistic:

```{r}
# Group Pb concentrations by location_code (DL or YD), calculate grouped means
soil_loc_means <- soil |> 
    summarize(pb_ppm_mean = mean(pb_ppm),
                .by = location_code) |> 
    arrange(location_code)

# Calculate diff_means (DL - YD)
diff_means = soil_loc_means$pb_ppm_mean[1] - soil_loc_means$pb_ppm_mean[2]
```

## Shuffle the predictor variable and recalculate the statistic

```{r}
# Shuffle predictor variable (location_code)
shuffle_soil <- soil |> 
    mutate(shuffle_location_code = sample(location_code)) |>
    summarize(pb_ppm_mean = mean(pb_ppm),
                .by = shuffle_location_code) |> 
    arrange(shuffle_location_code)

# Calculate diff_means (DL - YD)
shuffle_diff_means <- shuffle_soil$pb_ppm_mean[1] - shuffle_soil$pb_ppm_mean[2]

```

## Repeat Shuffling

```{r}
# Replicate above shuffle 10K times (1e4)
shuffled_means <- replicate(1e4,{
                    # Shuffle predictor variable (location_code)
                    shuffle_soil <- soil |> 
                        mutate(shuffle_location_code = sample(location_code)) |>
                        summarize(pb_ppm_mean = mean(pb_ppm),
                                    .by = shuffle_location_code) |> 
                        arrange(shuffle_location_code)

                    # Calculate diff_means (DL - YD)
                    shuffle_soil$pb_ppm_mean[1] - shuffle_soil$pb_ppm_mean[2]
                    })

#shuffled_means
```

```{r}
# Plotting histogram (null distribution)
tibble(shuffled_means) |> 
    ggplot(aes(shuffled_means)) + 
    geom_histogram(binwidth = 5)

```

## Compare the observed to the null

```{r}
tibble(diff_means = shuffled_means) |>
  ggplot(aes(diff_means)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = diff_means, color = "magenta", linewidth = 1.5)

```

```{r}
# Calculate pvalue
pval <- mean(shuffled_means > diff_means)

pval
```

## Interpret your results

We would reject the null hypothesis since the pval < 0.5 (threshold).

# Claim 2: Neighborhoods north of the Exide plant were exposed to more lead than those south.

## 1. Calculate observed value of the statistic

```{r}
ena_esa <- soil |>
  mutate(area = str_sub(property_id, 1, 3)) |>
  filter(area %in% c("ENA", "ESA"))

area_summ <- ena_esa |>
  summarize(mean_pb_ppm = mean(pb_ppm), .by = area) |>
  arrange(area)

obs_diff_means <- area_summ$mean_pb_ppm[1] - area_summ$mean_pb_ppm[2]

obs_diff_means
```

## 2. Resample observations and recalculate the statistic

```{r}
# Resample rows from the data frame and recalculate the statistic.
# Repeat this 10,000 times to simulate the sampling distribution.
area_samp_dist <- replicate(
  1e3,
  {
    resampled_summ <- ena_esa |>
      slice_sample(n = nrow(ena_esa), replace = TRUE) |>
      summarize(mean_pb_ppm = mean(pb_ppm), .by = area) |>
      arrange(area)

    resampled_summ$mean_pb_ppm[1] -
      resampled_summ$mean_pb_ppm[2]
  }
)
```

```{r}
# Visualize the sampling distribution using a histogram
tibble(diff_means = area_samp_dist) |>
  ggplot(aes(diff_means)) +
  geom_histogram(binwidth = 5)
```

## 3. Find the confidence interval

A confidence interval is a range of values that you are confident contains the population parameter.

```{r}
# Use the quantile() function to find the middle 95% of the sampling distribution.
area_95ci <- quantile(area_samp_dist, c(0.025, 0.975))
```

```{r}
# Plot the sampling distribution again, but this time add vertical lines indicating the observed statistic and the 95% CI.

tibble(diff_means = area_samp_dist) |>
  ggplot(aes(diff_means)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = obs_diff_means, color = "magenta", linewidth = 1.5) +
  geom_vline(xintercept = area_95ci, linetype = "dashed", linewidth = 1.5)
```

## 4. Interpret your results

The difference in Pb exposure between north and south could be 40-100ppm.

This means the north generally did have more PB exposure than the south (positive 40-100ppm 95% of the time).